---
title: "Habitat Classification and Resistance Surfaces using Unmanned Aerial Vehicles"
author: "Ben Weinstein"
date: "July 2, 2015"
output:
  html_document:
    keep_md: yes
    toc: true
---

```{r,message=FALSE,warning=FALSE}
library(raster)
library(maptools)
library(ggplot2)
library(plyr)
library(reshape2)
library(rgdal)
library(plotKML)
library(rgeos)
library(SDMTools)
library(rasterVis)

#set dropbox path
droppath<-"C:/Users/Ben/Dropbox/"

#load("Drone.RData")
```

#Preprocessing
* The data was collected using a Phantom Vision 2+ drone that was shot in manual flight. The photos were shot in RAW and batch transformed to .tif in Photoshop.


* An orthomosiac was created in Agisoft which produced a .tif of three bands, and a Digital Elevation Model (DEM).

#Goals

* To read in the points
* Classify the habitat using kmeans unsupervised classifcation
* Use the ground truth data to check classification accuracy.

##Read in GPS points

```{r}
#Georeferenced points relating to the classes
pt<-readGPX(paste(droppath,"Droning/PamGPSExtract20150701_2.gpx",sep="/"))
head(pt$waypoints)
```

The key here is the lon, lat, ele and name, which will match the waypoint ID

## Read in point classification

Pam was in the field and recorded the habitat class of each point

```{r}
cl<-read.csv(paste(droppath,"Droning/Whetstone_groundtruth_landscapefeatures3.csv",sep="/"))

#merge with the gpx data
pts<-merge(x=pt$waypoints,y=cl,by.x="name",by.y="Waypoint")
```

How many points in each class?

```{r}
table(pts$Landscape.class)
#Turn into spatial class
pts<-SpatialPointsDataFrame(cbind(pts$lon,pts$lat),pts)
projection(pts)<-CRS("+proj=longlat +datum=WGS84")
```

##Read in the spectral raster 
crop by extent of points
```{r}
r1<-brick("C:/Users/Ben/Dropbox/Droning/20150425_Run6.tif")
```

reproject points to UTM
```{r}
#rpts<-spTransform(pts,CRS(projection(r1)))
rpts<-pts
e<-extent(rpts)
```

Crop raster to points
```{r,eval=T}
plot(extent(r1))
plot(extent(e),col='red',add=T)
rcrop<-crop(x=r1,y=e*1.01)
```

##read in DEM

```{r,eval=T}
#clip by the spatial extent of the points.
d<-raster(paste(droppath,"Droning/20150425_Run6_DEM.tif",sep="/"))

#reproject?
#dpr<-projectRaster(d,rcrop)

#crop
dcrop<-crop(d,extent(rcrop))
names(dcrop)<-c("DEM")

#remove negative elevation?
dcrop[dcrop<0]<-0
```

View DEM

```{r}
plot(dcrop)
```

###Overlay points 

```{r,fig.height=10,fig.width=10}
plot(rcrop)
points(pts,col=pts$Landscape.class,pch=16,cex=1.5)
```

##Extract points from the raster

```{r}
names(rcrop)<-c("Band1","Band2","Band3","Band4")
vals<-extract(x=rcrop[[1:3]],y=pts,sp=T,buffer=2,fun=mean)
```

##Unsupervised classifcation

The points are not aligning to our expectation of the classes. Let's have the computer decided the best differentation of the top 4 classes and then assign names to each?

Unsupervised classication is done through kmeans clustering, where we supply the known number of clusters. K-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster. The process is done iteratively, starting with random centers and comparing the fit of each solution.

Give the relative hetergenity of the system, we can aggregate cells and make this slightly less computationally exhausting. 

##SPECTRAL
```{r,eval=T}
ragg<-aggregate(rcrop,20)
```

##DEM

Make DEM into the same dimension as the aggregated spectral.

```{r,eval=T}
fact<-res(ragg)/res(dcrop)
#fact=50
#round to 3 decimnal places, we will need to enforce the same size
dagg<-aggregate(dcrop,fact=fact)
plot(dagg)

#there can be rounding errors, enforce the same extent
extent(dagg)<-extent(ragg)
```

#Create habitat classification
##Two Stage process - DEM then Spectral

Use the dem to classify the trees and the bushes, but then use the spectral to classify the swale and hummock. To do this we need everything to have the same spatial resolution
* Classify into 3 the dem aggregate raster

```{r}
#3 classes
#dagg[is.na(dagg[])]<-0

cl<-kmeans(x=dagg[],centers=3,iter.max=100,nstart=3)
kmeansDagg<-raster(dagg)
kmeansDagg[]<-cl$cluster
plot(kmeansDagg,col=topo.colors(10))
points(rpts,col=rpts$Landscape.class,pch=16,cex=.8)
text(rpts,labels=rpts$Landscape.class,pch=16,cex=.7)
```

Get the portions of the aggregated spectral raster, where the classified aggregated dem raster is not bushes or trees.

```{r}
#hard code a value in there, ordinal scale changes.
swale<-extract(kmeansDagg,cbind(-122.9093,42.42011))

#make a copy
kmeancopy<-kmeansDagg
plot(tomask<-Which(!kmeancopy==swale))
kmeancopy[tomask]<-NA
spectralnobush<-mask(ragg,kmeancopy)
plot(spectralnobush)
```

Classify the resulting raster, between types.

```{r}
spectralnobush[is.na(spectralnobush)]<-0
cl<-kmeans(x=spectralnobush[[1:3]][],centers=3,iter.max=100,nstart=3)
kmeansspectralnobush<-raster(spectralnobush)
kmeansspectralnobush[]<-cl$cluster
plot(kmeansspectralnobush,col=topo.colors(10))
```

Combine the spectral class with the dem classification

```{r}
combo<-kmeansspectralnobush* kmeansDagg
unique(combo)
plot(combo)
levelplot(combo,margin=F,par.settings=RdBuTheme())
```

##Overlay on image

```{r}
plotRGB(ragg)
plot(combo,alpha=.3,add=T,axes=F)
```

```{r}
#extract value from fixed locations
tree<-extract(combo,cbind(-122.9083,42.41852))
bush<-extract(combo,cbind(-122.9096,42.42054))
swale<-extract(combo,cbind(-122.9093,42.4202))
hummock<-extract(combo,cbind(-122.908,42.4198))
```

**Tree is :`r tree`**
**Bush is :`r bush`**
**Swale is :`r swale`**
**Hummock is :`r hummock`**

Assign classes

```{r}
combo <- ratify(combo)
rat <- levels(combo)[[1]]

#create lookup table
lookup<-data.frame(value=c(swale,hummock,bush,tree),cl=c('swale','hummock','bush','tree'))

rat<-merge(rat,lookup,by.x="ID",by.y="value",all=T)

#remove NA
combo[combo==rat[is.na(rat$cl),"ID"]]<-NA

#repeat without NA, probably not the cleanest way to do that
combo <- ratify(combo)
rat <- levels(combo)[[1]]

#create lookup table
lookup<-data.frame(value=c(swale,hummock,bush,tree),cl=c('swale','hummock','bush','tree'))

rat<-merge(rat,lookup,by.x="ID",by.y="value",all=T)

levels(combo) <- rat

levelplot(combo,col.regions=c("deeppink4","goldenrod2","dodgerblue4","darkgreen"))

#save as driver?
#jpeg("HabitatMap.jpg",res=300,units="in",height=7,width=9)
#levelplot(combo,col.regions=c("deeppink4","goldenrod2","dodgerblue4","darkgreen"))
#dev.off()

```

### Landscape ecology metrics

We can measure fragmentation, size, area, shape, etc. For example for each of our classes (bush,swale,hummock,tree = 1,2,3,4)

```{r,fig.height=20,fig.width=15}
#
stats<-ClassStat(combo,res(combo)[1])

#merge with rat table
stats<-merge(stats,rat,by.x="class",by.y="ID")

stats<-melt(stats,id.vars="cl")
ggplot(stats,aes(x=cl,y=value)) + geom_bar(stat="identity") + facet_wrap(~variable,scales="free",ncol=4)
```

```{r}
save.image("Drone.RData")
```
