---
title: "Habitat Classification using Unmanned Aerial Vehicles"
author: "Ben Weinstein"
date: "July 2, 2015"
output:
  html_document:
    keep_md: yes
---

```{r,message=FALSE,warning=FALSE}
library(raster)
library(maptools)
library(ggplot2)
library(plyr)
library(reshape2)
library(rgdal)
library(plotKML)

#set dropbox path
droppath<-"C:/Users/Ben/Dropbox/"
```

#Preprocessing
* The data was collected using a Phantom Vision 2+ drone that was shot in manual flight. The photos were shot in RAW and batch transformed to .tif in Photoshop.

* An orthomosiac was created in Agisoft which produced a .tif of three bands, and a DEM. The origin of the DEM remains a mystery to me.

#Goals

* To read in the points
* Classify the habitat using kmeans unsupervised classifcation
* Use the ground truth data to check classification accuracy.

##Read in GPS points

```{r}
#Georeferenced points relating to the classes
pt<-readGPX(paste(droppath,"Droning/PamGPSExtract20150701_2.gpx",sep="/"))
head(pt$waypoints)
```

The key here is the lon, lat, ele and name, which will match the waypoint ID

## Read in point classification

Pam was in the field and recorded the habitat class of each point

```{r}
cl<-read.csv(paste(droppath,"Droning/Whetstone_groundtruth_landscapefeatures3.csv",sep="/"))

#merge with the gpx data
pts<-merge(x=pt$waypoints,y=cl,by.x="name",by.y="Waypoint")
```

How many points in each class?

```{r}
table(pts$Landscape.class)
#Turn into spatial class
pts<-SpatialPointsDataFrame(cbind(pts$lon,pts$lat),pts)
```

Draw extent of points

```{r}
e<-extent(pts)
```

##Read in the spectral raster 
crop by extent of points
```{r}
r<-brick(paste(droppath,"Droning/AgisoftFull20150425_Run3.tif",sep="/"))
rcrop<-crop(r,e*1.1)
plot(rcrop)
```

##read in DEM

```{r}
#clip by the spatial extent of the points.
d<-raster(paste(droppath,"Droning/AgisoftFull20150425_Run3_DEM.tif",sep="/"))
dcrop<-crop(d,e*1.1)
names(dcrop)<-c("DEM")
plot(dcrop)
```

###Overlay points 
As you can see, the points don't quite match yet. Brendan and Pam are checking on the georeferencing of the drone layer. We can continue with the habitat classifcation and wait to test the accuracy of the results.

```{r,fig.height=10,fig.width=10}
plot(rcrop)
points(pts,col=pts$Landscape.class,pch=16,cex=1.5)
```

##Extract points from the raster

```{r}
names(rcrop)<-c("Band1","Band2","Band3","Band4")
vals<-extract(x=rcrop[[1:3]],y=pts,sp=T,buffer=5,fun=mean)
```

Let's learn about the values

```{r}
mvals<-melt(vals@data,measure.vars=c("Band1","Band2","Band3"))
ggplot(mvals,aes(x=Landscape.class,y=value,fill=variable)) + geom_point() + theme_bw()
```

Do we need to resample the dem to fit the resolution of the spectral?

##Unsupervised classifcation

The points are not aligning to our expectation of the classes. Let's have the computer decided the best differentation of the top 4 classes and then assign names to each?

Unsupervised classication is done through kmeans clustering, where we supply the known number of clusters. K-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster. The process is done iteratively, starting with random centers and comparing the fit of each solution.

Give the relative hetergenity of the system, we can aggregate cells and make this slightly less computationally exhausting. 

##SPECTRAL
```{r}
ragg<-aggregate(rcrop,50)
```

##DEM

Make DEM into the same dimension as the aggregated spectral.

```{r}
fact<-res(ragg)/res(dcrop)
#round to 3 decimnal places, we will need to enforce the same size
dagg<-aggregate(dcrop,fact=fact)
plot(dagg)
```

##Two Stage process

Use the dem to classify the trees and the bushes, but then use the spectral to classify the swale and hummock. To do this we need everything to have the same spatial resolution

* Classify into 3 the dem aggregate raster

Make DEM into the same dimension as the aggregated

```{r}
#3 classes
cl<-kmeans(x=dagg[],centers=3,iter.max=10,nstart=3)
kmeansDagg<-raster(dagg)
kmeansDagg[]<-cl$cluster
plot(kmeansDagg,col=topo.colors(10))
points(pts,col=pts$Landscape.class,pch=16,cex=.8)
text(pts,labels=pts$Landscape.class,pch=16,cex=.7)
```

Get the portions of the aggregated spectral raster, where the classified aggregated dem raster is not bushes or trees.

```{r}
kmeansDagg[!kmeansDagg==1]<-NA
spectralnobush<-mask(ragg,kmeansDagg)
plot(spectralnobush)
```

Classify the resulting raster, between types.

```{r}
spectralnobush[is.na(spectralnobush)]<-0
cl<-kmeans(x=spectralnobush[],centers=3,iter.max=10,nstart=3)
kmeansspectralnobush<-raster(spectralnobush)
kmeansspectralnobush[]<-cl$cluster
plot(kmeansspectralnobush,col=topo.colors(10))
```

Combine the spectral class with the dem classification

```{r}
combo<-kmeansspectralnobush* kmeansDagg
unique(combo)
```

1 is Bush
2 is Tree
6 is Hummock
9 is Swale

The remaining question is whether the delinaiation of swale is accurate and fits out expectation surrounding the bushes. If look at the rim of the bushes, you see a kind of swale halo that matches the classification of the river of swale. we need to be sure that fits monica's expectation for the genetic sampling so we have the same terminology!

#Computing geographic distances
#Computing resistance distance
#Computing least cost path